Analyzing the Effect of Surface in Tennis Grand Slams

### 2018-09-13 contest/surface-type/paper-tennis.Rmd
added gs partial point by point data.  3066 out of 5000+ matches

# Introduction {#sec:iintro}
Rafael Nadal is known as the "King of Clay" in tennis, having won 11 out of his current 17 grand slams titles at the French Open, which is played on a clay surface [@bbc2018].  In contrast, his rival Roger Federer has won his most grand slam titles (8 out of 20) at Wimbledon, which is played on grass.  On the women's side, Serena Williams has been dominant both on hard court (7 titles at the Australian Open and 6 at US Open) and grass (7 at Wimbledon).  This trend extends to other top players, who seem to have better results at some grand slams than others.  More broadly, it seems that country of origin has an interaction effect with court type.  For example, Spanish players seem to excel on clay courts and Americans have great success at Wimbledon despite grass courts not being of wide use in the USA.  It also worth questioning whether the US Open and Australian Open should be grouped together as hard courts despite having different surface compositions [@theage2007].  In this paper, we analyze the results of grand slam players from 2013-2017,  and we

@@

For issue (2) the player analysis of Nadal, Federer, and Williams, we examine whether our model passes the "common sense" tests  like how the models in [thomas2013] show that commonly well known hockey players also have high status in the model.  We also examine whether these players do have surface apparent effects. Few academic papers have been written about Nadal, Federer, or Williams.  One paper studies Federer's odds of winning when Nadal suddenly withdrew from Wimbledon showed that Federer was too heavily favored by bookmakers [@leitner2009].  One analysis of Williams shows how she has gotten better with age, even past the point when other greats began to decline, but the study does not look at surface type [@five2015].

Finally, for issue (3), we use clustering methods [cite] in order to determine which court surface types are more similar to one another.

@@

## Data {#sec:data}
The data consists of `r nrow(gs)` matches split evenly over the four grand slams and the two leagues: ATP (men's) and WTA (women's).  Each match has `r ncol(gs)` attributes, many of which are redundant.  We focus on the following attributes for both the winner and loser of the match: games won, points won, retirement, break points faced, break points saved, aces, country of origin, and player attributes.  Additionally, we take into account the number of sets in a match, the surface type, minutes played, and round of the tournament.  A subset of the data is shown in Table \ref{tab:data}.

@@

The data is obtained from Jeff Sackman's open website [cite] via the `R` package `deuce` [cite].  All steps of our analysis from collection to dissemination are freely available online and are compiled into an `R` package \url{http://github.com/shannong19/courtsports}.

@@

>>>>

# Introduction {#sec:iintro}
Rafael Nadal is known as the "King of Clay" in tennis, having won 11 out of his current 17 grand slams titles at the French Open, which is played on a clay surface [@bbc2018].  In contrast, his rival Roger Federer has won his most grand slam titles (8 out of 20) at Wimbledon, which is played on grass.  On the women's side, Serena Williams, winner of 24 grand slam titles, has been dominant both on hard court (7 titles at the Australian Open and 6 at US Open) and grass (7 at Wimbledon).  This trend extends to other top players, who seem to have better results at some grand slams than others.  More broadly, it seems that country of origin has an interaction effect with court type.  For example, Spanish players seem to excel on clay courts and Americans have great success at Wimbledon despite grass courts not being of wide use in the USA.  It also worth questioning whether the US Open and Australian Open should be grouped together as hard courts despite having different surface compositions [@theage2007].  In this paper, we analyze the results of grand slam players from 2013-2017,  and we

@@

For issue (2) the player analysis of Nadal, Federer, and Williams, we examine whether our model passes the "common sense" tests  like how the models in [@thomas2013] show that commonly well known hockey players also have high status in the model.  We also examine whether these players do have surface apparent effects. Few academic papers have been written about Nadal, Federer, or Williams.  One paper studies Federer's odds of winning when Nadal suddenly withdrew from Wimbledon showed that Federer was too heavily favored by bookmakers [@leitner2009].  One analysis of Williams shows how she has gotten better with age, even past the point when other greats began to decline, but the study does not look at surface type [@five2015].

Finally, for issue (3), we use clustering methods in order to determine which court surface types are more similar to one another.

@@

## Data {#sec:data}
The data consists of `r nrow(gs)` matches split evenly over the four grand slams and the two leagues: ATP (men's) and WTA (women's).  Each match has `r ncol(gs)` attributes, many of which are redundant.  We focus on the following attributes for both the winner and loser of the match: games won, points won, retirement, break points faced, break points saved, aces, country of origin, and player attributes.  Additionally, we take into account the number of sets in a match, the surface type, and round of the tournament.  A subset of the data is shown in Table \ref{tab:data}.

@@

The data is obtained from Jeff Sackmann's open website via the `R` package `deuce` [@sackmann2018; @deuce2017].  All steps of our analysis from collection to dissemination are freely available online.

@@

\footnotesize

### 2018-09-13 contest/surface-type/paper-tennis.tex
added gs partial point by point data.  3066 out of 5000+ matches

Rafael Nadal is known as the ``King of Clay'' in tennis, having won 11
out of his current 17 grand slams titles at the French Open, which is
played on a clay surface (Jurejko 2018). In contrast, his rival Roger
Federer has won his most grand slam titles (8 out of 20) at Wimbledon,
which is played on grass. On the women's side, Serena Williams has been
dominant both on hard court (7 titles at the Australian Open and 6 at US
Open) and grass (7 at Wimbledon). This trend extends to other top
players, who seem to have better results at some grand slams than
others. More broadly, it seems that country of origin has an interaction
effect with court type. For example, Spanish players seem to excel on
clay courts and Americans have great success at Wimbledon despite grass
courts not being of wide use in the USA. It also worth questioning
whether the US Open and Australian Open should be grouped together as
hard courts despite having different surface compositions (Paxinos
2007). In this paper, we analyze the results of grand slam players from
2013-2017, and we

@@

For issue (2) the player analysis of Nadal, Federer, and Williams, we
examine whether our model passes the ``common sense'' tests like how the
models in {[}thomas2013{]} show that commonly well known hockey players
also have high status in the model. We also examine whether these
players do have surface apparent effects. Few academic papers have been
written about Nadal, Federer, or Williams. One paper studies Federer's
odds of winning when Nadal suddenly withdrew from Wimbledon showed that
Federer was too heavily favored by bookmakers (Leitner, Zeileis, and
Hornik 2009). One analysis of Williams shows how she has gotten better
with age, even past the point when other greats began to decline, but
the study does not look at surface type (Morris 2015).

Finally, for issue (3), we use clustering methods {[}cite{]} in order to
determine which court surface types are more similar to one another.

@@

The data is obtained from Jeff Sackman's open website {[}cite{]} via the
\texttt{R} package \texttt{deuce} {[}cite{]}. All steps of our analysis
from collection to dissemination are freely available online and are
compiled into an \texttt{R} package
\url{http://github.com/shannong19/courtsports}.

@@

\hypertarget{sec:refs}{%
\section*{References}\label{sec:refs}}
\addcontentsline{toc}{section}{References}

@@

@@

>>>>

Rafael Nadal is known as the ``King of Clay'' in tennis, having won 11
out of his current 17 grand slams titles at the French Open, which is
played on a clay surface (Jurejko 2018). In contrast, his rival Roger
Federer has won his most grand slam titles (8 out of 20) at Wimbledon,
which is played on grass. On the women's side, Serena Williams, winner
of 24 grand slam titles, has been dominant both on hard court (7 titles
at the Australian Open and 6 at US Open) and grass (7 at Wimbledon).
This trend extends to other top players, who seem to have better results
at some grand slams than others. More broadly, it seems that country of
origin has an interaction effect with court type. For example, Spanish
players seem to excel on clay courts and Americans have great success at
Wimbledon despite grass courts not being of wide use in the USA. It also
worth questioning whether the US Open and Australian Open should be
grouped together as hard courts despite having different surface
compositions (Paxinos 2007). In this paper, we analyze the results of
grand slam players from 2013-2017, and we

@@

For issue (2) the player analysis of Nadal, Federer, and Williams, we
examine whether our model passes the ``common sense'' tests like how the
models in (Thomas et al. 2013) show that commonly well known hockey
players also have high status in the model. We also examine whether
these players do have surface apparent effects. Few academic papers have
been written about Nadal, Federer, or Williams. One paper studies
Federer's odds of winning when Nadal suddenly withdrew from Wimbledon
showed that Federer was too heavily favored by bookmakers (Leitner,
Zeileis, and Hornik 2009). One analysis of Williams shows how she has
gotten better with age, even past the point when other greats began to
decline, but the study does not look at surface type (Morris 2015).

Finally, for issue (3), we use clustering methods in order to determine
which court surface types are more similar to one another.

@@

The data is obtained from Jeff Sackmann's open website via the
\texttt{R} package \texttt{deuce} (Sackmann 2018; Kovalchik 2017). All
steps of our analysis from collection to dissemination are freely
available online.

@@

\hypertarget{sec:refs}{%
\section{References}\label{sec:refs}}

\footnotesize

@@

\leavevmode\hypertarget{ref-deuce2017}{}%
Kovalchik, Stephanie. 2017. \emph{Deuce: Resources for Analysis of
Professional Tennis Data}.

@@

\leavevmode\hypertarget{ref-sackmann2018}{}%
Sackmann, Jeff. 2018. ``Tennis Data Repositories.''
\url{https://github.com/JeffSackmann}.

\leavevmode\hypertarget{ref-thomas2013}{}%
Thomas, A. C., Samuel L. Ventura, Shane T. Jensen, and Stephen Ma. 2013.
``COMPETING Process Hazard Function Models for Player Ratings in Ice
Hockey.'' \emph{The Annals of Applied Statistics} 7 (3). Institute of
Mathematical Statistics: 1497--1524.
\url{http://www.jstor.org/stable/23566482}.
